Hardware-Accelerated Computation of Stellar Light Curves

### 2024-08-09 latex/paper.tex
feat: add all error benchmark tex and scripts

        \caption{Error of the solution vector $\bvec{s}$ and the overall flux $f$ for all terms of the spherical harmonics basis up to degree $l=20$. Errors are computed against arbitrary-precision computations of $\bvec{s}$ and $f$ for a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$. For each $(l, m)$ basis component, errors are scaled to the highest values of $\bvec{s}$ (respectively $f$) over $b$. Then, the plotted errors for each degree $l$ are the maximum of the scaled errors over all $m\in [-l, l]$ components over $b$. The disks at the top of the figure show the transit configurations of the occultor (light dark) and the star (light gray) for different values of $b$. \codelink{workflows/precision/scripts/plot_error}}

@@

@@

        \includegraphics[width=\textwidth]{../workflows/speed/figures/error_rb.png}
        \caption{relative numerical error in the flux $f$ of a linearly limb-darkened star transited by an opaque companion over the $(r, b)$ parameter space. The error is evaluated for different orders $n$ of the Gauss-Legendre quadrature and computed against \texttt{starry} for speed and convenience. Note that \autoref{fig:precision_s} was produced using $n=500$ (right plot) for which $\bvec{s}$ reaches machine precision.}

@@

At a fixed occultor radius, \autoref{fig:relative_error_2} shows the time required to compute the flux of the system for a range of $N$ impact parameters $b$, plotted against the maximum relative error reached, the order $n$ of the Gauss-Legendre approximation, and the hardware used for the computation.

% \begin{figure}[H]
%     \begin{center}
%         \includegraphics[width=\textwidth]{../workflows/performance/figures/error_vs_time.pdf}
%         \caption{}
%         \label{fig:relative_error_2}
%     \end{center}
% \end{figure}
These computations were run on an apple M1 CPU and an NVIDIA ... GPU. Thanks to the use of JAX, no harware-specific... \\\\

By computing $\mathcal{P}$ numerically, we show that our vectorized implementation reaches machine-precision on a GPU, up to an order of magnitude faster than the current C++ implementation of \texttt{starry} ran on a CPU.\\\\

Plot of the precision versus speed of the computed light curves.

Plot of the solution vectors computation similar to figure 11 and 12 of luger (maybe in the appendix)

@@

>>>>

        \caption{Error of the solution vector $\bvec{s}$ and the overall flux $f$ for all terms of the spherical harmonics basis up to degree $l=20$. Errors are computed against arbitrary-precision computations of $\bvec{s}$ and $f$ for a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$. For each $(l, m)$ basis component all coefficients are set to zero except $y_{l,m}=1$ and errors are scaled to the highest values of $\bvec{s}$ (respectively $f$) over $b$. Then, the plotted errors for each degree $l$ are the maximum of the scaled errors over all $m\in [-l, l]$ components over $b$. The disks at the top of the figure show the transit configurations of the occultor (light dark) and the star (light gray) for different values of $b$. \codelink{workflows/precision/scripts/plot_error}}

@@

For reference, \autoref{fig:precision_limbdark} shows the relative error in the flux of a limb-darkened star occulted by an opaque body for increasing orders of a polynomial limb-darkening law. In this figure, we also report the error of the flux computed using the \textsf{exoplanet} Python package (limited to linear and quadratic limb-darkening).
        \includegraphics[width=\textwidth]{../workflows/precision/figures/limbdark_error.PDF}
        \caption{Maximum errors in the overall flux $f$ of a star occulted by an opaque body for increasing orders of its polynomial limb-darkening law. Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, taken over the range of values $b$. For each order $n$ all limb-darkening coefficients are set to zero except $u_n=1$. \codelink{workflows/precision/scripts/plot_error_limbdark}}
        \label{fig:precision_limbdark}
A set of unitary polynomial limb-darkening coefficients of order 20 leads to spherical harmonics coefficients as large as $10^6$. This explains the larger errors observed in \autoref{fig:precision_limbdark} (error versus limb-darkening order) compared to the ones shown in \autoref{fig:precision_degree} (error versus degree of the spherical harmonics). We not that the errors scale with the values of $u$.\\\\

@@

        \includegraphics[width=\textwidth]{../workflows/precision/figures/br_error.png}
        \caption{relative numerical error in the flux $f$ of a linearly limb-darkened star (degree $l$=1) transited by an opaque companion over the $(r, b)$ parameter space. The error is evaluated for different orders $n$ of the Gauss-Legendre quadrature and computed against \texttt{starry} for speed and convenience. Note that \autoref{fig:precision_s} was produced using $n=500$ (right plot) for which $\bvec{s}$ reaches machine precision. \codelink{workflows/precision/scripts/plot_br_error}}

@@

In \autoref{fig:error_order_gausslegendre}, we show how the error in the solution vector evolves with the order $n$ of the Gauss-Legendre quadrature, and its impact on the error in the overall flux.
        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_order.pdf}
        \caption{Maximum errors in the solution vector and overall flux $f$ for different occultor radii and orders $n$ of the Gauss-Legendre quadrature used to compute the P integral (see \autoref{solution_vectors}). Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$. As in \autoref{fig:precision_s} the maximum degree of the spherical harmonics basis is set to $l=20$.}
        \label{fig:error_order_gausslegendre}
The minimum error is always achieved for higher values of $n$, which comes with higher computational cost (see \autoref{speed}). Hence users must carefully set the order parameter to balance precision and computational speed. As the minimum error reached also depends on the degree of the spherical harmonics basis used, \autoref{fig:error_order_gausslegendre_degree} shows the errors on the flux for each degree of the spherical harmonics basis, for an occultor with relative radii $r=0.01$, $r=1$ and $r=100$.
        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_order_degree.pdf}
        \caption{Maximum errors in the overall flux $f$ for all degrees of the spherical harmonics components for different occultor radii and orders $n$ of the Gauss-Legendre quadrature used to compute the P integral (see \autoref{solution_vectors}). Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$.}
        \label{fig:error_order_gausslegendre_degree}

@@

The main advantage of the \texttt{jax} library is its ability to perform hardware-accelerated computations on CPUs and GPUs. In this section, we evaluate the speed of the \texttt{jaxoplanet} implementation and compare it with the C++ implementation of \texttt{starry}. For quadratically limb-darkened light curve, we also compare our implementation with the one provided by the \textsf{exoplanet} Python package.\\\\
In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the computing time required to compute an occultation light-curve for a limb-darkened star and a non-uniform star described by spherical harmonics up to order $l=20$. This figure include the value of the order $n$ of the Gauss-Legendre quadrature required to reach a given precision, and illustrate well the trade-off between precision and speed that users must consider when setting the order parameter.\\\\
\begin{figure}[H]
    \begin{center}
        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_order.pdf}
        \caption{Computation time of the occultation light curve of a quadts}
        \label{fig:speed_order_gausslegendre}
    \end{center}
\end{figure}

\todo{} shows
